Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-12 Final Draft/Final Draft.Rmd
updates

@@

```
\FloatBarrier
This coefficient is still somewhat difficult to interpret, and made increasingly so by the winner-take-all system by which U.S. Electoral College votes are allocated. As an alternative perspective, we examined a range of counterfactual scenarios, in which counties experienced job losses
<!--  what should I read in the counterfactual -->

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

#  make map
flipped <- mapping_small %>% filter(changed_outcome == "Yes")
ggplot() +
  geom_sf(data = mapping_small, aes(fill = delta_vs_group), color = "black", size = .1) +
  scale_fill_manual(values = colors) +
  labs(title = "",
       subtitle = "",
       fill = "Change in Democratic Vote Share") +
  geom_sf(data = flipped, fill = NA, 
          aes(color = changed_outcome), 
          size = 1) +
  scale_color_manual(values = "#b32ab5", name = "Flipped Election") +
  labs(color = "Flipped Election") +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank())

@@

\caption{Clinton Electoral College Votes}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

>>>>

Like \cite{Baccini21}, we find evidence that the experience of deindustralization was associated with support for Donald Trump, particularly among whites. \cite{Baccini21} find that gross manufacturing layoffs between 2012-2015 were associated with a loss in Democratic vote share. However, using a different measure of deindustrialization - *net* change in manufacturing employment - we find the oppposite effect: that counties with manufacturing job *gains* from 2012-2015 tended to swing towards Donald Trump in 2016. We extend the analysis to the 2004-2015 period in order to capture the long-term effects of deindustrialization on a region. Doing so, we find that regions experiencing more manufacturing job losses swung towards Donald Trump in 2016, and that this occurred more when those losing jobs were white.

@@

```
\FloatBarrier
This coefficient is still somewhat difficult to interpret, and made increasingly so by the winner-take-all system by which U.S. Electoral College votes are allocated. As an alternative perspective, we examined a range of counterfactual scenarios, in which counties manufacturing experienced job losses between the 1st and 99th percentiles across all counties.

For example, in the map below, we assume that all counties in the US experienced a loss in white manufacturing employment equivalent to the level experienced by a county at the 25th percentile. Here, that level is equivalent to a 0 net loss in white manufacturing jobs between 2004-2015 per 100 workers in 2004 (a relatively low level of job losses). Setting all counties' white manufacturing job losses to 0, we use the point estimates of our regression coefficients from Model 3 in Table \ref{regResult04} to estimate the counterfactual change in Democratic vote share. We find that, had all counties experienced the 25th percentile level in manufacturing job losses, most states would have leaned more Democratic in 2016, and that Michigan and Pennsylvania would have flipped.

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

#  make map
flipped <- mapping_small %>% filter(changed_outcome == "Yes")
ggplot() +
  geom_sf(data = mapping_small, aes(fill = delta_vs_group), color = "black", size = .1) +
  scale_fill_manual(values = colors) +
  labs(title = "",
       subtitle = "",
       fill = "Change in Democratic Vote Share") +
  geom_sf(data = flipped, fill = NA, 
          aes(color = changed_outcome), 
          size = 1) +
  scale_color_manual(values = "#b32ab5", name = "Flipped Election") +
  labs(color = "Flipped Election") +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.key.size = unit(.3, 'cm'))

@@

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

```
\begin{figurenotes}
Figures show a 95% confidence interval for Democratic Electoral College Votes at Each Percentile of Manufacturing Job Losses.
\end{figurenotes}
\end{figure}
\FloatBarrier

### 2021-12-12 Final Draft/Final-Draft.tex
updates

In this choice we diverge from \cite{Baccini21}, which uses
\textbf{gross} manufacturing job losses. For example, if a county lost
\(450\) manufacturing jobs from 2012-2015 and also gained \(400\)
manufacturing jobs over the same period, this would be \(450\)
\textbf{gross} job losses, but only \(50\) \textbf{net} job losses. The
\textbf{gross} measure captures several dynamics unrelated to
deindustrialization - for example, seasonal unemployment in a
food-manufacturing region, or workers moving between jobs. Thus we
believe the \textbf{net} job losses over a period are the more accurate
measure. It can be argued that the non-seasonal or non-structural
components of gross job losses are a useful covariate. Basic notions of
prospect theory suggest that the loss of a job and subsequent regaining
of another identical job may still create substantial discontent.
Constructing a better measure of gross job losses is an area for further
methodological and substantive research.

@@

The results can be found in Table \ref{regResult04}. The regression
coefficient in Model 4, for instance, means that a loss of \(10\)
manufacturing jobs per \(100\) workers in a county (a relatively extreme
value) corresponds to a loss of \(3.2 \%\) in Democratic vote share.
\FloatBarrier

@@

This coefficient is still somewhat difficult to interpret, and made
increasingly so by the winner-take-all system by which U.S. Electoral
College votes are allocated. As an alternative perspective, we examined
a range of counterfactual scenarios, in which counties experienced job
losses

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

@@

\caption{Clinton Electoral College Votes}

@@

>>>>

In this choice we diverge from \cite{Baccini21}, which uses \emph{gross}
manufacturing job losses. For example, if a county lost \(450\)
manufacturing jobs from 2012-2015 and also gained \(400\) manufacturing
jobs over the same period, this would be \(450\) \emph{gross} job
losses, but only \(50\) \emph{net} job losses. The \emph{gross} measure
captures several dynamics unrelated to deindustrialization - for
example, seasonal unemployment in a food-manufacturing region, or
workers moving between jobs. Thus we believe the \emph{net} job losses
over a period are the more accurate measure. Additionally, there is
precedent in the literature to use \emph{net} employment changes, as
\cite{Autor21} have done. Nevertheless, it can be argued that the
non-seasonal or non-structural components of gross job losses are a
useful covariate. For instance, basic notions of prospect theory suggest
that the loss of a job and subsequent regaining of another identical job
may still create substantial discontent. Constructing a better measure
of gross job losses is an area for further methodological and
substantive research.

@@

The results can be found in Table \ref{regResult04}. The regression
coefficient in Model 4, for instance, means that a loss of \(10\) white
manufacturing jobs per \(100\) workers in a county (a relatively extreme
value) corresponds to a \(3.2\) percentage point loss in Democratic vote
share. \FloatBarrier

@@

This coefficient is still somewhat difficult to interpret, and made
increasingly so by the winner-take-all system by which U.S. Electoral
College votes are allocated. As an alternative perspective, we examined
a range of counterfactual scenarios, in which counties manufacturing
experienced job losses between the 1st and 99th percentiles across all
counties.

For example, in the map below, we assume that all counties in the US
experienced a loss in white manufacturing employment equivalent to the
level experienced by a county at the 25th percentile. Here, that level
is equivalent to a 0 net loss in white manufacturing jobs between
2004-2015 per 100 workers in 2004 (a relatively low level of job
losses). Setting all counties' white manufacturing job losses to 0, we
use the point estimates of our regression coefficients from Model 3 in
Table \ref{regResult04} to estimate the counterfactual change in
Democratic vote share. We find that, had all counties experienced the
25th percentile level in manufacturing job losses, most states would
have leaned more Democratic in 2016, and that Michigan and Pennsylvania
would have flipped.

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

@@

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }

@@

\FloatBarrier
\end{figure}
\begin{figurenotes}
Figures show a 95% confidence interval for Democratic Electoral College Votes at Each Percentile of Manufacturing Job Losses.
\end{figurenotes}
\FloatBarrier
